Spin canonical rings of log stacky curves

### 2015-07-02 stacky_write_up/spin_cring.tex
added remark of precise statement of main corollary

The main theorem of this paper, Theorem ~\ref{thm:main}, combined
with a lemma of Voight and Zureick-Brown \cite[Lemma 10.2.1]
{vzb:stacky}, provides a statement on rings of modular forms.
Together with the main theorem from Voight and Zureick-Brown \cite
[Theorem 9.3.1]{vzb:stacky}, we have the following application to
rings of modular forms. Note that the lemma from Voight and Zureick-
Brown ~\cite[Lemma 10.2.1]{vzb:stacky} follows almost immediately
from the classical work of Kodaira--Spencer
\cite{kodaira-spencer:deformations}.

@@

@@

\begin{enumerate}
\item[(a)] $R_{\halfcan}$ is generated over $R_{\halfcan'}$ by 
	elements $y_1, \ldots, y_n$ such that $\deg(y_i) = d_i$ and $-\ord_P
	^{L'}(y_i) = c_i$.

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
	\[
		\ord_u(f) < \ord_u(h) \implies f\prec h.
	\]
	Equip $k[y_1, \ldots, y_n]$ with graded $P$-lex, as defined in
	Definition ~\ref{defn:graded-p-lex} and $k[u, x_1, \ldots, x_m, y_1,
	\ldots, y_n]$ with the block order giving the order giving $y_1, 
	\ldots, y_n$ precedence over $u, x_1, \ldots, x_m$.

	If $I'$ and $I$ are the ideals of relations of $k[u, x_1, \ldots, x_m]
	\to R_{\halfcan'}$ and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]
	\to R_{\halfcan}$ respectively, then

	\begin{align*}
		\initial(I) &= \initial(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
											 + \langle U_i: 1 \le i \le n-1 \rangle
											 + \langle V \rangle,\\
	\end{align*}
	where $V = \{x_i y_j: 1 \le i \le m, 1 \le j \le n\}$ and $U_i$ is
	the set of monomials of the form $\prod_{j = 1}^{i} y_j^{a_j}$ with
	$a_j \in \BN_{\ge 0}$ such that
	\begin{enumerate}
		\item[(*)] $\sum_{j=1}^i a_j c_j \ge c_{i+1}$, \\
		\item[(**)] there does not exist $(b_1, \ldots b_i) \ne (a_1,
			\ldots a_i)$ with all $b_j \le a_j$ and $\sum_{j = 1}^i b_j
			c_j \ge	c_{i + 1}$, \\
		\item[(***)] there does not exist $r<i$ such that $\sum_{j=1}^r y_j^{a_j}\ge c_{r + 1}$.
	\end{enumerate}

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$
	in degrees at most $\beta$ with $I$ generated over $I'$ in
	degrees at most $\max(2 \beta, \beta + \tau)$ (where $\tau = \max(1,
	\max_{1 \leq i \leq m}(\deg(x_i)))$).
\end{enumerate}
\end{lem}

@@

\todo{how to avoid polynomials hangining over lines like this, ask david} Since the ideal in $k[u, y_1,\ldots, y_n]$ generated by is $S$ is $(y_1, \ldots, y_n) \cdot k[u, y_1, \ldots, y_n]$ and $\bigcup_{i=1}^{n-1} U_i$ generates every monomial in $\cup_{i=1}^n y_i \cdot k[u, y_1, \ldots, y_n]-S$, we have shown that all monomials of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ are contained in
\begin{align*}
				& k[u, x_1, \ldots, x_m] \cup \langle V \rangle \cup S \cup \left\langle \bigcup_{i=1}^{n-1} U_i \right\rangle \\
	\subseteq \; 	& S \cup \langle T\rangle \cup \initial(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; k[u, x_1, \ldots, x_m].
\end{align*}
This shows condition (3) of Lemma ~\ref{lem:relations_from_generators_induction} holds.  Thus, the conditions of 
Lemma ~\ref{lem:relations_from_generators_induction} are met. Finally, Lemma ~\ref{lem:relations_from_generators_induction} implies part (b).  \todo{Peter: Checking that all the conditions of Lemma \ref{lem:relations_from_generators_induction}(2) are met takes a lot of space and clutter.  On the other hand, it is somewhat important.  Should it be included?}

@@

{\bf Part (c):}
To show (c), we wish to show that $y_{i, e_i}z \in R'$ for all generators $z$ of $R'$
with $z \neq y_{i, e_i} \text{ and } z \neq y_{i, e_i'}$. Note that $f \in R$ further satisfies $f \in R'$ if and
only if for all $j \in J$ we have
\begin{align}
\label{eqn:order-degree}
	-\ord_{Q_j}
^{\halfcan_X'}(f) \leq \deg (f) \left( \subhalf {e_j'} \right).
\end{align}
\todo{Peter: I understand why this should be true, although its not clear to me that there it easily follows.  Maybe we should explain it a little as to not confuse the reader. Aaron: Can we ask David, this was clear to me in his paper, and he also didn't include it.}

@@

%\begin{lem}
%\label{lem:genus-2-generation-5}
%Let $X$ be a smooth genus 2 curve with canonical divisor $K$ and let $L \in \di X$ be so that $2L \sim K$. 
%Then, $R_\halfcan$ is generated in degree at most $5$.
%\end{lem}
%\begin{proof}
%By Lemma ~\ref{lem:semicanonical_generation} it suffices to show $R_\halfcan$ has no minimal generators in degree 6.
%Let $P_1, \ldots, P_6$ be the six hyperelliptic fixed points of $X$. 
%Recall that all linear equivalence classes of spin divisors on $X$ have a representative of the form
%$$L = P_1 + \sum_{i =2}^{6} \epsilon_i (P_1 - P_i),$$ 
%with $\epsilon_i \in \{0,1\},$ and not all $\epsilon_i = 1$. \todo{cite this somewhere, maybe ask david for a place to cite? We can cite http://www.math.harvard.edu/~ctm/home/text/class/harvard/213b/14/html/home/course/course.pdf, for example, from complex analysis course notes! Don't use coarse notes}
%Note that $H^0(X,K)$ is spanned by rational functions $u, x$ with $\di u|_{2L} = -2L, \di x|_{2L} = -2L + 2P_1,$ where $u,x$ are though of as rational functions. Therefore, the image of the multiplication map 
%\begin{align*}
	%Sym^3 H^0(X,2L) \rightarrow H^0(X,6L)
%\end{align*}
%is 4 dimensional by Riemann--Roch, with image spanned by $u^3,u^2x,ux^2, x^3$, 
%which have 
%\begin{align*}
	%\di u^3|_{2L} &= -2L \\
	%\di u^2x|_{2L} &= -2L + 2P_1 \\
	%\di ux^2|_{2L} &= -2L+4P_1, \\
	%\di x^3|_{2L} &= -2L + 6P_1.
%\end{align*}
%
%\noindent
%Since $H^0(X,6L)$ is 5 dimensional by Riemann--Roch,
%it suffices to produce one more independent function in the image of the multiplication map.
%There are now two cases, depending on whether $H^0(X,L) = 0$.
%
%First, we deal with the case that $H^0(X,L) \neq 0$. 
%Then, since $X \not \cong \BP^1_k,$ and $\deg L = 1$, we must have $H^0(X,L) = 1$. Hence, we know $L$ is effective, so up to linear equivalence, we may assume $L = P$ where $P$ is hyperelliptic fixed. In this case, it is shown in \cite{neves:halfcan}[Theorem III.19] that $R_\halfcan$ is generated in degree $\leq 3$.
%
%Second, suppose $H^0(X,L) = 0$.
%Note that because $\deg 3L = 3$, we have by Riemann--Roch that $H^0(X,3L) = 2$. 
%Note additionally that $3L \sim K + L \sim 2P_1 + L$. Since $3L$ is base point free, by Riemann--Roch
%we have that there exists a rational function $f$ with $\di f|_{2P_1 + L} = -2P_1 - L.$ Additionally, there are no functions $g \in H^0(X,L + 2P_1)$ 
%with $\di g|_{2P_1 + L} = -L,$ since $H^0(X,L) = 0.$ Therefore, there must 
%exist a function $h \in H^0(X,L+2P_1)$ so that $\di h|_{L + 2P_1} = -L - P_1$.
%Then, $h \cdot f \in H^0(X,2L + 4P_1) \cong H^0(X,6L)$ has a pole of odd 
%order at $P_1$, and is therefore independent from the image of $\sym^3 H^0(X,3L)$.
%\end{proof}

>>>>

The main theorem of this paper, Theorem ~\ref{thm:main}, combined
with a lemma of Voight and Zureick-Brown \cite[Lemma 10.2.1]
{vzb:stacky}, provides a statement on rings of modular forms.
Together with the main theorems from Voight and Zureick-Brown \cite
[Theorem 9.3.1]{vzb:stacky} and \cite
[Theorem 9.3.1]{vzb:stacky}, we have the following application to
rings of modular forms. Note that the lemma from Voight and Zureick-
Brown ~\cite[Lemma 10.2.1]{vzb:stacky} follows almost immediately
from the classical work of Kodaira--Spencer
\cite{kodaira-spencer:deformations}.

@@

\begin{rem}
In the case that $M_k(\Gamma) = 0$ for all odd $k$, the generation bound of $3 \cdot \max(3, e_1, \ldots, e_r)$ and relation
bound of $6 \cdot \max(3, e_1,\ldots, e_r)$ can be reduced to $\max(3,e_1, \ldots, e_r)$ and $2 \cdot \max(3,e_1, \ldots, e_r)$, apart from several small families of cases. See 
~\cite[Theorem 9.3.1]{vzb:stacky} and ~\cite[Theorem 9.3.1]{vzb:stacky} for a more precise statement of these bounds.
\end{rem}

@@

\begin{enumerate}
\item[(a)] $R_{\halfcan}$ is generated over $R_{\halfcan'}$ by elements $y_1, \ldots, y_n$ such that $\deg(y_i)=d_i$ and $-\ord_P^{L'}(y_i)=c_i$.

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
\[
	\ord_u(f) < \ord_u(h) \implies f\prec h.
\]
Equip $k[y_1, \ldots, y_n]$ with graded $P$-lex, as defined in Definition ~\ref{defn:graded-p-lex} and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ with the block order giving the order giving $y_1, \ldots, y_n$ precedence over $u, x_1, \ldots, x_m$.  
If $I'$ and $I$ are the ideals of relations of $k[u, x_1, \ldots, x_m]\to R_{\halfcan'}$ and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]\to R_{\halfcan}$ respectively, then

\begin{align*}
	\initial(I) &= \initial(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle,\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_j^{a_j}$ with $a_j \in \BN_{\ge 0}$ such that
\begin{enumerate}
	\item[(*)] $\sum_{j=1}^i a_j c_j \ge c_{i+1},$\\
	\item[(**)] there does not exist $(b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ with all $b_j\le a_j$ and $\sum_{j=1}^i b_jc_j\ge c_{i+1},$ \\
	\item[(***)] there does not exist $r<i$ such that $\sum_{j=1}^r y_j^{a_j}\ge c_{r+1}$.
\end{enumerate}

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with $I$ generated over $I'$ in degrees at most $\max(2\beta, \beta + \tau)$ (where $\tau = \max(1, \max_{1\le i, m}(\deg(x_i)))$).
\end{enumerate}
\end{lem}

@@

Since the ideal in $k[u, y_1,\ldots, y_n]$ generated by is $S$ is $(y_1, \ldots, y_n) \cdot k[u, y_1, \ldots, y_n]$ and $\bigcup_{i=1}^{n-1} U_i$ generates every monomial in $\cup_{i=1}^n y_i \cdot k[u, y_1, \ldots, y_n]-S$, we have shown that all monomials of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ are contained in
\begin{align*}
				& k[u, x_1, \ldots, x_m] \cup \langle V \rangle \cup S \cup \left\langle \bigcup_{i=1}^{n-1} U_i \right\rangle \\
	\subseteq \; 	& S \cup \langle T\rangle \cup \initial(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; k[u, x_1, \ldots, x_m].
\end{align*}
This shows condition (3) of Lemma ~\ref{lem:relations_from_generators_induction} holds.  Thus, the conditions of 
Lemma ~\ref{lem:relations_from_generators_induction} are met. Finally, Lemma ~\ref{lem:relations_from_generators_induction} implies part (b).  \todo{Peter: Checking that all the conditions of Lemma \ref{lem:relations_from_generators_induction}(2) are met takes a lot of space and clutter.  On the other hand, it is somewhat important.  Should it be included?}

@@

{\bf Part (c):}
To show (c), we wish to show that $y_{i, e_i}z \in R'$ for all generators $z$ of $R'$
with $z \neq y_{i, e_i} \text{ and } z \neq y_{i, e_i'}$. Note that $f \in R$ further satisfies $f \in R'$ if and
only if for all $j \in J$ we have
\begin{align}
\label{eqn:order-degree}
	-\ord_{Q_j}
^{\halfcan_X'}(f) \leq \deg (f) \left( \subhalf {e_j'} \right).
\end{align}

\todo{Peter: I understand why this should be true, although its not clear to me that there it easily follows.  Maybe we should explain it a little as to not confuse the reader. Aaron: Can we ask David, this was clear to me in his paper, and he also didn't include it. Say by the definition}

@@

%\begin{lem}
%\label{lem:genus-2-generation-5}
%Let $X$ be a smooth genus 2 curve with canonical divisor $K$ and let $L \in \di X$ be so that $2L \sim K$. 
%Then, $R_\halfcan$ is generated in degree at most $5$.
%\end{lem}
%\begin{proof}
%By Lemma ~\ref{lem:semicanonical_generation} it suffices to show $R_\halfcan$ has no minimal generators in degree 6.
%Let $P_1, \ldots, P_6$ be the six hyperelliptic fixed points of $X$. 
%Recall that all linear equivalence classes of spin divisors on $X$ have a representative of the form
%$$L = P_1 + \sum_{i =2}^{6} \epsilon_i (P_1 - P_i),$$ 
%with $\epsilon_i \in \{0,1\},$ and not all $\epsilon_i = 1$.
%Note that $H^0(X,K)$ is spanned by rational functions $u, x$ with $\di u|_{2L} = -2L, \di x|_{2L} = -2L + 2P_1,$ where $u,x$ are though of as rational functions. Therefore, the image of the multiplication map 
%\begin{align*}
	%Sym^3 H^0(X,2L) \rightarrow H^0(X,6L)
%\end{align*}
%is 4 dimensional by Riemann--Roch, with image spanned by $u^3,u^2x,ux^2, x^3$, 
%which have 
%\begin{align*}
	%\di u^3|_{2L} &= -2L \\
	%\di u^2x|_{2L} &= -2L + 2P_1 \\
	%\di ux^2|_{2L} &= -2L+4P_1, \\
	%\di x^3|_{2L} &= -2L + 6P_1.
%\end{align*}
%
%\noindent
%Since $H^0(X,6L)$ is 5 dimensional by Riemann--Roch,
%it suffices to produce one more independent function in the image of the multiplication map.
%There are now two cases, depending on whether $H^0(X,L) = 0$.
%
%First, we deal with the case that $H^0(X,L) \neq 0$. 
%Then, since $X \not \cong \BP^1_k,$ and $\deg L = 1$, we must have $H^0(X,L) = 1$. Hence, we know $L$ is effective, so up to linear equivalence, we may assume $L = P$ where $P$ is hyperelliptic fixed. In this case, it is shown in \cite{neves:halfcan}[Theorem III.19] that $R_\halfcan$ is generated in degree $\leq 3$.
%
%Second, suppose $H^0(X,L) = 0$.
%Note that because $\deg 3L = 3$, we have by Riemann--Roch that $H^0(X,3L) = 2$. 
%Note additionally that $3L \sim K + L \sim 2P_1 + L$. Since $3L$ is base point free, by Riemann--Roch
%we have that there exists a rational function $f$ with $\di f|_{2P_1 + L} = -2P_1 - L.$ Additionally, there are no functions $g \in H^0(X,L + 2P_1)$ 
%with $\di g|_{2P_1 + L} = -L,$ since $H^0(X,L) = 0.$ Therefore, there must 
%exist a function $h \in H^0(X,L+2P_1)$ so that $\di h|_{L + 2P_1} = -L - P_1$.
%Then, $h \cdot f \in H^0(X,2L + 4P_1) \cong H^0(X,6L)$ has a pole of odd 
%order at $P_1$, and is therefore independent from the image of $\sym^3 H^0(X,3L)$.
%\end{proof}
